---
title: "Lab 4"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

## 1. Read in the data

```{r packages}
library(data.table)
library(tidyverse)
```

```{r data-read, cache=TRUE}
if (!file.exists("../met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "../met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../met_all.gz")
```

## 2. Prepare the data

```{r data-prep, cache=TRUE}
# Remove temperatures less than -17C
met <- met[temp >= -17]

# Make sure there are no missing data in the key variables coded as 9999, 999, etc
# temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met[, range(temp)]
met[, range(rh, na.rm = TRUE)]
met[, range(wind.sp, na.rm = TRUE)]
met[, range(vis.dist, na.rm = TRUE)]
met[, range(dew.point, na.rm = TRUE)]
met[, range(lat, na.rm = TRUE)]
met[, range(lon, na.rm = TRUE)]
met[, range(elev, na.rm = TRUE)]

met[elev == 9999.0, elev := NA]

# Generate a date variable using the functions as.Date()
# (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
met[, ymd := as.Date(paste(year, month, day, sep = "-"))]

# Using the data.table::week function, keep the observations of
# the first week of the month.
met[, table(week(ymd))]
met <- met[ week(ymd) == 31]

# Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met_avg <- met[, .(
  temp      = mean(temp, na.rm = TRUE),
  rh        = mean(rh, na.rm = TRUE),
  wind.sp   = mean(wind.sp, na.rm = TRUE),
  vis.dist  = mean(vis.dist, na.rm = TRUE),
  dew.point = mean(dew.point, na.rm = TRUE),
  lat       = mean(lat, na.rm = TRUE),
  lon       = mean(lon, na.rm = TRUE),
  elev      = mean(elev, na.rm = TRUE), USAFID
), by = "USAFID"]

# Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71  & lon <= -98, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98 , region := "Northeast"]
met_avg[lat < 39.71  & lon > -98 , region := "Southeast"]

met_avg[, region2 := fifelse(lat >= 39.71 & lon <= -98, "Northwest",
        fifelse(lat < 39.71  & lon <= -98, "Southwest",
                fifelse(lat >= 39.71 & lon > -98, "Northeast",
                        fifelse(lat < 39.71  & lon > -98, "Southeast", NA_character_))))]

met_avg[, table(region, region2, useNA = "always")]
# Create a categorical variable for elevation as in the lecture slides
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]

# Deleting extra column
met_avg <- met_avg[, -1]
```

## 3. Use geom_violin to examine the wind speed and dew point temperature by region

```{r violin-temp}
ggplot(met_avg, mapping = aes(y = wind.sp, x = 1)) +
  geom_violin() +
  facet_grid(~region)
```

## 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region

## 5

## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r plot-stat_summ}
ggplot(
  met_avg[!is.na(wind.sp) & !is.na(dew.point)],
  mapping = aes(x = region, y = wind.sp)) +
  stat_summary(fun.data = "mean_sdl") +
  stat_summary(fun.data = "mean_sdl", geom = "errorbar")
```

## 7. Make a map showing the spatial trend in relative h in the US

```{r leaflet}
library(leaflet)

temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)

leaflet(met_avg) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(round(rh,2), ' rh'), color = ~ temp.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=temp.pal, values=met_avg$rh,
          title='Temperature, C', opacity=1)
  
```

